Headgroup Structure and Cation Binding in Phosphatidylserine Lipid Bilayers

### 2019-09-03 Manuscript/manuscriptPS.tex
Add mention about Han NBfix in the conclusions

First, we complemented the available experimental information \cite{browning80,roux90} by measuring the signs of the PS headgroup order parameters,
and then proceeded to compare MD simulation results from several force fields with the experimental data.
This revealed that none of the force fields
reproduce the PS headgroup order parameters within the experimental accuracy.
However, the best models captured essential differences between PS and PC, and
suggested characteristic conformations of PS headgroups.
Comparison to the experimentally observed order parameters in POPC:POPS (5:1) bilayers  at varying ion
concentrations \cite{roux90} then showed that the tested MD force fields
overestimate the cation binding affinity to these bilayers. There were two exceptions: 1) the MacRog force field with potassium counterions, which appeared to display a more realistic monovalent ion binding
affinity to PS-containing lipid bilayers than the other models; and 2) the CHARMM36 force field with the recently introduced
NBfix correction for calcium \cite{kim16}, which underestimated the calcium binding affinity.
Importantly, the experimentally measured responses of the PS headgroup $S_\mathrm{CH}$ to bound calcium, as well as to dilution of the bilayer with zwitterionic PC lipids, were not
qualitatively reproduced in any of the tested force fields.
This is in contrast to previous results with PC lipids,
where MD force fields were seen to be at least in qualitative agreement with the experimentally measured headgroup $S_\mathrm{CH}$ responses to bound charge---even
though the headgroup structures themselves were
incorrect and the cation binding affinities were overestimated \cite{catte16}.
This underlines the dire need for more realistic MD force fields to study the biological function of PS lipids.

>>>>

First, we complemented the available experimental information \cite{browning80,roux90} by measuring the signs of the PS headgroup order parameters,
and then proceeded to compare MD simulation results from several force fields with the experimental data.
This revealed that none of the force fields
reproduce the PS headgroup order parameters within the experimental accuracy.
However, the best models captured essential differences between PS and PC, and
suggested characteristic conformations of PS headgroups.
Comparison to the experimentally observed order parameters in POPC:POPS (5:1) bilayers  at varying ion
concentrations \cite{roux90} then showed that the tested MD force fields
overestimate the cation binding affinity to these bilayers. There were two exceptions: 1) the MacRog force field with potassium counterions, which appeared to display a more realistic monovalent ion binding
affinity to PS-containing lipid bilayers than the other models; and 2) the CHARMM36 force field with the recently introduced
NBfix corrections for calcium \cite{kim16,han2018graph}, which underestimated the calcium binding affinity.
Importantly, the experimentally measured responses of the PS headgroup $S_\mathrm{CH}$ to bound calcium, as well as to dilution of the bilayer with zwitterionic PC lipids, were not
qualitatively reproduced in any of the tested force fields.
This is in contrast to previous results with PC lipids,
where MD force fields were seen to be at least in qualitative agreement with the experimentally measured headgroup $S_\mathrm{CH}$ responses to bound charge---even
though the headgroup structures themselves were
incorrect and the cation binding affinities were overestimated \cite{catte16}.
This underlines the dire need for more realistic MD force fields to study the biological function of PS lipids.
